### 2024-06-16 research-methods.qmd
20240615 - relabel
20240615 - reliability, validity

## Measurement Validity {#sec-measurementValidity}

### Reliability {#sec-reliability}

### Validity {#sec-validity}

>>>>

## Psychometrics {#sec-psychometrics}

### Measurement Reliability {#sec-reliability}

The *reliability* of a measure's scores deals with the *consistency* of measurement.
This book focuses on the following types of reliability:

- [test–retest reliability](#sec-testRetestReliability)
- [inter-rater reliability](#sec-interraterReliability)
- [intra-rater reliability](#sec-intraraterReliability)
- [internal consistency](#sec-internalConsistency)
- [parallel-forms reliability](#sec-parallelFormsReliability)

#### Test–Retest Reliability {#sec-testRetestReliability}

Test–retest reliability evaluates the consistency of scores across time.
For a construct that is expected to be stable across time (e.g., hand size in adults), we would expect our measurements to be consistent across time.
The consistency of scores across time can be examined in terms of relative or absolute test–retest reliability.
Relative test–retest reliability—i.e., the consistency of individual differences across time—is commonly evaluated using the coefficient of stability (i.e., the Pearson correlation coefficent).
Absolute test–retest reliability—i.e., the absolute consistency of people's scores across time—is commonly evaluated using the coefficient of repeatability.

#### Inter-Rater Reliability {#sec-interraterReliability}

Inter-rater reliability evaluates the consistency of scores across raters.
For instance, if we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player regardless of which (trained) rater (e.g., coach) uses it to rate the player.
The consistency of scores across raters is commonly evaluated using the intraclass correlation coefficient (for continuous variables) and Cohen's kappa ($\kappa$; for categorical variables).

#### Intra-Rater Reliability {#sec-intraraterReliability}

Intra-rater reliability evaluates the consistency of scores within a given rater.
If we have a strong measure for assessing college players' aptitude to succeed in the NFL, the measure should yield a similar score for a given player from the same (trained) rater (e.g., coach) each time they rate the same player.
The consistency of scores within raters can be evaluated using similar approaches as those evaluating [inter-rater reliability](#sec-interraterReliability).

#### Internal Consistency {#sec-internalConsistency}

Internal consistency evaluates the consistency of scores across items within a measure.
If we develop a strong questionnaire measure to assess a college players' aptitude to succeed in the NFL, the scores should be relatively consistent across items.
The consistency of scores across items within a measure is commonly evaluated using Cronbach's alpha ($\alpha$) or McDonald's omega ($\omega$).

#### Parallel-Forms Reliability {#sec-parallelFormsReliability}

Parallel-forms reliability evaluates the consistency of scores across different but equivalent forms of a measure.
If we develop two equivalent versions of the Wonderlic Contemporary Cognitive Ability Test (Form A and Form B) so that players sitting next to each other do not receive the same items, we would expect a player's score on Form A would be similar to their score on Form B.
Parallel-forms reliability is is commonly evaluated using the coefficient of equivalence (i.e., the Pearson correlation coefficent).

### Measurement Validity {#sec-validity}

The *validity* of a measure's scores deals with the *accuracy* of measurement.
This book focuses on the following types of validity:

- [face validity](#sec-faceValidity)
- [content validity](#sec-contentValidity)
- [criterion-related validity](#sec-criterionValidity)
    - [concurrent (criterion-related) validity](#sec-concurrentValidity)
    - [predictive (criterion-related) validity](#sec-predictiveValidity)
- [construct validity](#sec-constructValidity)
- [convergent validity](#sec-convergentValidity)
- [discriminant validity](#sec-discriminantValidity)
- [incremental validity](#sec-incrementalValidity)

#### Face Validity {#sec-faceValidity}

Face validity evaluates the extent to which a measure "looks like" (on its face) it assesses the construct of interest.
For instance, if a measure is developed to assess aptitude of Wide Receivers for the position, it would be considered to have face validity if everyday (lay) people believe that it assesses aptitude for being a successful Wide Receiver.

#### Content Validity {#sec-contentValidity}

Content validity evaluates the extent to which the measure assesses the full breadth of the content, as determined by context experts.
For the measure to have content validity, it should not have gaps (missing content facets) or intrusions (facets of other constructs).
For instance, a strong measure for assessing a player's aptitude to succeed in the NFL might need to include a player's speed, strength, size, lateral quickness, etc.
If the measure is missing their speed, this would be a content gap.
If the measure assesses a construct-irrelevant facet (e.g., their attractiveness), this would be a content intrusion.

#### Criterion-Related Validity {#sec-criterionValidity}

Criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest.
Criterion-related validity is commonly evaluated using a Pearson correlation or some form of regression.

There are two types of criterion-related validity:

- concurrent (criterion-related) validity
- predictive (criterion-related) validity

##### Concurrent (Criterion-Related) Validity {#sec-concurrentValidity}

Concurrent criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest assessed at the same point in time.
That is, concurrent validity could evaluate whether current player statistics (e.g., passing yards) are associated with their fantasy points.

##### Predictive (Criterion-Related) Validity {#sec-predictiveValidity}

Predictive criterion-related validity evaluates the extent to which the measure's scores are related to meaningful variables of interest that are assessed at a later point in time.
For example, predictive validity could evaluate whether scores on the measure we developed to assess a player's aptitude to succeed in the NFL predicts later performance in the NFL.

#### Construct Validity {#sec-constructValidity}

Construct validity evaluates the extent to which the measure's scores accurately assess the construct of interest.
If we develop a measure with intent to assess aptitude for being a successful Running Back, and it appears to more accurately assess aptitude for being a successful Wide Receiver, then our measure has poor construct validity for assessing aptitude for being a successful Running Back.
Construct validity subsumes [convergent](#sec-convergentValidity) and [discriminant](##sec-discriminantValidity) validity, in addition to all of the other forms of measurement validity.

#### Convergent Validity {#sec-convergentValidity}

Convergent validity evaluates the extent to which the measure's scores are related to other measures of the same construct.
For instance, if we develop a new measure to assess intelligence, its scores should be related to scores from other measures designed to assess intelligence (e.g., Wonderlic Contemporary Cognitive Ability Test).

#### Discriminant Validity {#sec-discriminantValidity}

Discriminant validity evaluates the extent to which the measure's scores are unrelated to measures of the different constructs.
For instance, if we develop a new measure to assess intelligence, its scores should be less strongly associated with measures of other constructs (e.g., measures of happiness).

#### Incremental Validity {#sec-incrementalValidity}

Incremental validity evalutes the extent to which the measure's scores provide an increase in predictive accuracy compared to other information that is easily and cheaply available.
That is, in order to be useful, a strong measure should tell us something that we did not already know.
For instance, if we develop a strong measure of intelligence, it should result in increased predictive accuracy (for success in the NFL) compared to when just relying on the Wonderlic Contemporary Cognitive Ability Test.

### Reliability vs Validity {#sec-reliabilityVsValidity}

Reliability and validity are different but related.
Reliability refers to the *consistency* of scores, whereas accuracy refers to the *accuracy* of scores.
Validity depends on reliability.
Reliability is necessary—but insufficient for—validity.
That is, consistency is necessary—but insufficient for—accuracy.
As depicted in @fig-reliabilityVsValidity, a measure can be no more valid than it is reliable.

::: {#fig-reliabilityVsValidity}
![](images/reliabilityVsValidity.png)

Reliability Versus Validity.
:::
